### 2025-03-21 about.qmd
Update bib

---
title: "Dirk Sliwka"
about:
  template: solana
  image: assets/img/SliwkaDirk.jpg
---

I am a professor of management in the Faculty of Management, Economics, and Social Sciences at the [University of Cologne](https://wiso.uni-koeln.de/en/) in the department [Corporate Development](https://cd.wiso.uni-koeln.de/en/). I am also senior research fellow at the [Max Planck Institute for Research on Collective Goods](https://www.coll.mpg.de/), a research fellow at [IZA](https://www.iza.org/person/790/dirk-sliwka) and [CESifo](https://www.cesifo.org/en), and a member of [C-SEB](https://c-seb.de/) and [ECONtribute](https://econtribute.de/).<br>
<br>
My research focuses on designing effective Human Resource Management practices within firms and evaluating their impact on behavior, performance, and well-being. I am particularly interested in the design of incentives and performance management processes as well as the role of social norms in organizations. <br>
<br>
My main research fields of interest are:<br>
<span style="display: block; margin-bottom: 6pt;"></span>
<ul>
    <li>Organizational/Personnel Economics</li>
    <li>Human Resource Management</li>
    <li>Field Experiments in Firms</li>
    <li>Behavioral Economics</li>
    <li>Management Accounting</li>
</ul>